---
title: "STAA 363 Lab1"
output: html_document
date: '2022-08-25'
---

##Question 1
```{r loading everything}
library(palmerpenguins)
library(ggplot2)
library(rlang)
suppressMessages(library(ggplot2))
```

```{rload data}
data("penguins")
```

Based on the information we have so far, we are dealing wtih regression modelling task.

##Question 2
For client 1, we are dealing with an association modeling task.

##Question 3
For client 2, we are dealing with a prediction modeling task.

```{r}
summary(penguins$body_mass_g)
histpenguine<-ggplot(penguins, mapping = aes(x=penguins$body_mass_g))+
  geom_histogram()
histpenguine
```

## Question 4
The smallest body mass in the dataset is 2700 grams.

## Question 5
50% of the penguines has body masss less than or equal to 4050 grams 

## Question 6

```{r count missing mass}
sum(is.na(penguins$body_mass_g))
```

There are 2 penguines in the data set that do not have a body mass recorded

```{r cleans}
penguinsClean<-na.omit(penguins)
```

## Question 7

```{r count missing}
nrow(penguins)
nrow(penguinsClean)
```

There are 11 rows that contained missing data in the penguins data set. There are 333 rows left with in the penguinsClean data set.

```{r histogram Penguine}
ggplot(penguinsClean, aes(x=body_mass_g)) + geom_histogram(bins = 20, fill='blue', col = 'black')

```
```{r}
ggplot(penguinsClean, aes(x=body_mass_g)) + geom_histogram(bins = 15, fill='blue', col = 'black')
```

