---
title: "Jason Hou STA 363 Lab 6"
author: "Jason Hou"
date: '2022-10-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
#Load all required packages
library(ggplot2)
library(broom)
library(gridExtra)
library(dplyr)
library(class)
library(tidyverse)
library(gridExtra)
library(leaps)
```

```{r load data}
#loading data needed
collegeData<-read.csv("collegedata.csv")
```

```{r take a look}
glimpse(collegeData)
```


## Question 1
*Is this a regression or classification task? Based on this, what metric will we likely use to assess predictive accuracy?*

It is a regression task since the response variable is numeric. We are likely to use the MSE/RMSE to assess the predictive accuracy

## Question 2
*Make a visualization to explore the distribution of graduation rate. What is the smallest value of graduation rate? The largest?*

```{r distribution}
ggplot(data=collegeData,aes(x=Grad.Rate))+
  geom_boxplot()+
  ggtitle("Figure 1. Distribution of Graduation Rate")
```

```{r max and min graduation rate}
gradmax<-max(collegeData$Grad.Rate)
gradmin<-min(collegeData$Grad.Rate)
knitr::kable(c("Largest" = gradmax,"Smallest" = gradmin),caption = "Largest and Smallest")
```



## Question 3

*Compute the MSE (training) and RMSE (training) for these data using Y^i=Y¯ for all rows i.*

```{r MSE and RMSE}
## MSE = RSS/n
MSE<-sum((collegeData$Grad.Rate-mean(collegeData$Grad.Rate))^2)/776
RMSE<-sqrt(MSE)
knitr::kable(c("MSE" = MSE,"RMSE" = RMSE),caption = "MSE and RMSE")
```

## Question 4
*Train an LSLR (OLS) regression model using the whole data set and the single feature X = the student faculty ratio. Find the MSE (training) and RMSE (training). How much has our training predictive accuracy improved from the values you got in Question 3?*

```{r}

```

