---
title: "Jason Hou STA 363 Lab 2"
output: html_document
date: '2022-09-01'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r load package}
library(tidyverse)
library(broom)
library(janitor)
library(ggplot2)
library(class)
suppressMessages(library(class))
```


```{r load data}
training<-read.csv("train.csv")
testing<-read.csv("test.csv")
```

```{r}
glimpse(training)
```

## Question 1

*How many rows are in the test data set? What about the training data set?*

```{r}
nrow(training)
```

```{r}
nrow(testing)
```

There are 900 rows in the test data set and 3012 rows in the testing data set.

## Question 2

*The test data set is smaller than the training data set. This is almost always the case in statistical learning applications. Explain why you think this is the case.*


## Question 3

*(a)Create a scatter plot where line breaks is on the x axis and the number of characters is on the Y axis.* 

*(b) Did you use your training or test data set to create your plot?*

```{r scatterplot}
Train_Scatter<-training%>%
  ggplot(mapping = aes(x=line_breaks,y=num_char))+
  geom_point(color="blue")+
  ggtitle("Figure 1: Scatterplot of Numer Line Breaks vs Number of Characteris in Email")
Train_Scatter
```

I used training data set to create the plot.

## Question 4

*Change the color and/or the shape of the points on the graph based on Y. In other words, spam emails should be one shape and/or color and non spam emails should be a different shape and/or color. Hint: Look back at Lab 1 if you have questions on how to do this!*

```{r change color shape}
Train_Scatter_withColor<-training%>%
  ggplot(mapping = aes(x=line_breaks,y=num_char,color=spam,pch=spam))+
  geom_point()+
  ggtitle("Figure 2: Scatterplot of Numer Line Breaks vs Number of Characteris in Email
          (by spam/not spam)")
Train_Scatter_withColor
```

## Question 5

*Suppose an email has 175 characters and 3800 line breaks. Using the 3-nearest neighbors, what value $\hat{Y}$ would you predict for this email? Hint: You don’t need any code for this; look at the graph from Question 4.*

Based on the graph from Question 4, we can locate this email near the top right corner of the graph, and I will predict  as not spam. 

## Question 6

*Using 5 nearest neighbors, run KNN. Look at the object results you have created. Is this a scalar (one value), a vector (one column or one row), or a matrix (multiple rows and columns)?*

```{r knn, include=FALSE}
results <- knn(train = training[,c(12,13)] , test=testing[,c(12,13)], cl = training$spam, k = 5 )
results
```

The object "result" is a vector.

```{r knn table}
table1<-table(results,testing$spam)
table1
```


## Question 7

*Make a table to show how many emails 5-nearest neighbors predicted to be spam and not spam.*

```{r table confusion Matrix}
knitr::kable(table(results, testing$spam), caption= "Table 1", col = c("True Not Spam", "True Spam"))
```

## Question 8

*What is the sensitivity of your classification method? Do we prefer methods with larger or smaller sensitivity?*

$\frac{18}{74+18}\approx0.19$

The sensitivity of the classification method is 0.19. We prefer methods to have a larer sensitivity.

## Question 9

*What is the specificity of your classification of your classification method? Do we prefer methods with larger or smaller specificity?*

$\frac{791}{791+17}\approx0.9789$

The specificity of the classification method is about 0.9785. We prefer method with higher specificity.


```{r}

```

## Question 10

*What is the false positive predictive rate (FPPR)? Do we prefer methods with larger or smaller FPPR?*

```{r}

```

## Question 11

*What is the false negative predictive rate (FNPR)? Do we prefer methods with larger or smaller FNPR?*


```{r}

```

## Question 12

*What is the accuracy of your classification approach? Hint: This the proportion of the predictions that are correct.*
