---
title: "STA 363 Lab 4"
author: "Jason Hou"
date: '2022-09-15'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(tidyverse)
library()
```


```{r load data and cleaning}
library(palmerpenguins)
data("penguins")

penguins <- na.omit(penguins)
```


## Question 1
*How many rows should be in each fold for our penguins data? Will all the folds have exactly the same number of rows? Explain your reasoning.*

```{r}
nrow(penguins)
```

We should have 33 rows in fold of the penguins data. Some folds will not have exactly the same number of rows since 333 is not fully divisible by 10.

## Question 2
*Create a data set called penguinsSmall that contains the first 20 rows in the penguin data set. If we use 5-fold CV with this smaller data set, how many rows belong in each fold?*

```{r small set}
penguinsSmall<-penguins[1:20,]
```

If we use 5-fold CV with this small data set, there will be 4 rows belong in each fold.

## Question 3

*For our 5-fold CV with penguinsSmall, how many slips of paper in this box need to contain the number 4?*

For the 5-fold CV with penguinSmall, there are 4 slips of paper in this box contain the number 4.

```{r}
papers <- rep(1:5, 4)
```

## Question 4

*Suppose we have a data set with 500 rows, and we want to use 5-fold CV. What code would need to create the equivalent of the papers vector for this new data set?*

```{r}
#Row500<-rep(1:100,5)
```


```{r}
set.seed(363663)
foldsSmall <- sample(papers, 20, replace = FALSE)
```

## Question 5
*Which fold is the 5th row in the data set assigned to? What about the 20th row?*

