---
title: "STA 363 Lab1"
output: html_document
date: '2022-08-25'
---

## Question 1
```{r loading everything}
library(palmerpenguins)
library(ggplot2)
library(rlang)
suppressMessages(library(ggplot2))
library(tidyverse)
library(gridExtra)
```

```{rload data}
data("penguins")
```

Based on the information we have so far, we are dealing wtih regression modelling task.

## Question 2

For client 1, we are dealing with an association modeling task.

## Question 3
For client 2, we are dealing with a prediction modeling task.

```{r}
summary(penguins$body_mass_g)
histpenguine<-ggplot(penguins, mapping = aes(x=penguins$body_mass_g))+
  geom_histogram()
histpenguine
```

## Question 4
The smallest body mass in the dataset is 2700 grams.

## Question 5
50% of the penguines has body masss less than or equal to 4050 grams 

## Question 6

```{r count missing mass}
sum(is.na(penguins$body_mass_g))
```

There are 2 penguines in the data set that do not have a body mass recorded

```{r cleans}
penguinsClean<-na.omit(penguins)
```

## Question 7

```{r count missing}
nrow(penguins)
nrow(penguinsClean)
```

There are 11 rows that contained missing data in the penguins data set. There are 333 rows left with in the penguinsClean data set.

```{r histogram Penguine}
ggplot(penguinsClean, aes(x=body_mass_g)) + geom_histogram(bins = 20, fill='blue', col = 'black')

```


```{r histogram penguine flipper length with bin 15}
ggplot(penguinsClean, aes(x=flipper_length_mm)) + geom_histogram(bins = 15, fill='cyan', col = 'white')
```

## Question 9

```{r w/title}
ggplot(penguinsClean, aes(x=flipper_length_mm)) + geom_histogram(bins = 15, fill='cyan', col = 'white')+
  labs(title = "Figure 2:",x="Flipper Length (in mm)",y="Frequency")
```

```{r boxplot}
MassBoxPlot<-penguinsClean%>%
  ggplot(mapping = aes(x=body_mass_g))+
  geom_boxplot(fill = "gold",col = "black")+
  labs(title = "Figure 3:",x="Body Mass(in grams)")
MassBoxPlot
```

## Question 11




```{r fillper length vs body mass scatter plot}
ggplot(penguinsClean, aes(x=flipper_length_mm, y = body_mass_g)) + geom_point()+labs(title = "Figure 4:",x = "Flipper Length(in mm)",y = "Body Mass (in grams)")

```

```{r scatter plot with LSLR line}
ggplot(penguinsClean, aes(x=flipper_length_mm, y = body_mass_g)) + geom_point()+labs(title = "Figure 5:",x = "Flipper Length(in mm)",y = "Body Mass (in grams)")+
stat_smooth(method = "lm", formula = y ~ x, size = 1, se = FALSE)
```

## Question 14
Yes, based on the shape of the relationship,it seems reasonable for LSLR to be used for the model.

```{r}
ggplot(penguinsClean, aes(body_mass_g)) + geom_histogram(bins = 20, fill = "blue", col = "white")

```

