---
title: "Jason Hou STA 363 Lab 7"
author: "Jason Hou"
date: '2022-11-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(palmerpenguins)
# Grows the tree
library(rpart)
# Allows us to visualize the tree
library(rattle)
library(rpart.plot)
library(dplyr)
```

```{r data}
data(penguins)
penguins <- na.omit(penguins)
```

## Question 1
*Is this an association task or a prediction task?*

This is an association task because the response variable is numeric.


## Question 2
*In the root, what is the predicted value $\hat{Y_i}$ for all penguins?*

```{r root model}
predicted<-mean(penguins$body_mass_g)
knitr::kable(predicted,caption = "Table 1: Predicted Value of Root Node")
```

The predicted value $\hat{Y_i}$ for all penguins in root is 4207.06.




## Question  3
*What is the RSS and RMSE with all the data in the root node (i.e, using no features)?*


```{r RMSE}
RMSE_root<-sqrt(mean((penguins$body_mass_g - predicted)^2))
RSS<-sum((penguins$body_mass_g-predicted)^2)
RMSE_RSS<-data.frame("RMSE" = RMSE_root,"RSS" = RSS)
knitr::kable(RMSE_RSS,caption = "Table 2: RMSE and RSS of Root Node")
```

## Quetion 4
*With species as a feature, what are all the possible splitting rules we could use to divide the root into two leaves? Note: Order doesn’t matter.*

```{r}
table(penguins$species)%>%
  knitr::kable(caption = "Table 3. Type of Peguin Species in Data Set")
```

We can see from  Table 3 that there are 3 types of penguins in the data set,therfore, there are 3 splitting rules we could use to divide the root into two leaves, they are:
1. Is the penguin an Adelie penguin?
2. Is the penguin an Chinstrap penguin?
3. Is the penguin an Gentoo penguin?

## Question 5

```{r simulate tree}
n<-nrow(penguins)
Y<-penguins$body_mass_g
X<-penguins$species

#TypeSpecies<-seq("Adelie","Chinstrap","Gentoo")
species<-c("Adelie","Chinstrap","Gentoo")
RSS<-rep(0,length(species))
t <- 1
for(i in species){
  leaf1<-which(X==i)
  leaf2<-c(1:n)[-leaf1]
  mean1<-mean(Y[leaf1])
  mean2<-mean(Y[leaf2])

  preds<-rep(0,n)
  
  preds[leaf1]<-mean1
  preds[leaf2]<-mean2
  RSS[t]<-sum((Y-preds)^2)
  t<-t+1
}
```



## Question 6

```{r}
species[which.min(RSS)]
```

Based on the result of Question 5, I recommend that we use "Is the penguin a Gentoo penguin" to split the root node into 2 trees. 


## Question 7

```{r}
tree1<- rpart(body_mass_g ~ species, data = penguins, method = "anova", maxdepth = 1)

# Option 1 
fancyRpartPlot(tree1, sub = "Figure 1: One Split")
```

We can see from Figure 1 that 36 percent of penguins are in leaf 2.

## Question 8

The tree predicts that all Chinstrap penguins has body mass of 3715 grams.

## Question 9