---
title: "STA 363 Project 1"
author: "Jason Hou"
date: '2022-09-12'
output: html_document
---

# Abstract



# Section 1 Introduction and Data

We are working on the data set that records 5726 children with potential dengue fever, which is a disease that is difficult to diagnose. The data set contains the basic information and measures of the children, including age, sex, BMI(Body Mass Index), height and weight, and 9 features about potential symptoms of dengue fever.The goal of our client is to use this data to establish a efficacious tool as an alternative to predict the dengue fever in children who lacks the access to standard diagnosis test for dengue fever.    

```{r load packages}
#Load all required packages
library(ggplot2)
library(broom)
library(gridExtra)
library(dplyr)
library(class)
library(tidyverse)
library(gridExtra)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load data}
#Load dengue data
dengue<-read.csv("dengueData.csv")
```

```{r checking}
#Checking if NA exists in data set
knitr::kable(sum(is.na(dengue)),col.names = "Missing Value")
```

No missing value was detected from the data set.


```{r look, include=FALSE}
#Take a look on data type and what the data looks like
glimpse(dengue)
```


#Section 2 KNN

## KNN Example

```{r data set for demo}
#Create a data set for demo
knnDemo<-data.frame("dengue_status"= dengue$Y,"Age"= dengue$Age, "Height"= dengue$Height)
```

```{r knn demo loop, eval=FALSE, include=FALSE}
#Number of rows in demo
n<-nrow(knnDemo)
#Set a seed for sampling
set.seed(363663)
demoSelectRow<-sample(1:n,n*0.9,replace = "FALSE")
demoTrain<-knnDemo[demoSelectRow,]
demoTest<-knnDemo[-demoSelectRow,]

true1<-which(demoTest$dengue_status == 1)
true0<-which(demoTest$dengue_status == 0)

ntrue1<-length(true1)
ntrue0<-length(true0)

storeK<-data.frame("K"=rep(NA,50),"GMean"=rep(NA,50))

for(i in 1:50){
  storeK$K[i]<-i
  Kpreds<-knn(demoTrain[,c(2,3)],demoTest[,c(2,3)],k=i,cl=demoTrain$dengue_status)
  sensitivity<-sum(Kpreds[true1] == 1)/ntrue1
  specificity<-sum(Kpreds[true0] == 0)/ntrue0
  storeK$GMean[i]<-sqrt(sensitivity*specificity)
}

knnDemoPlot<-ggplot(storeK,aes(K,GMean))+
  geom_line()+
  labs(caption = paste("Geometric Mean, ReD Line at K=", which.max(storeK$GMean)),title = "Figure 1",y = " ")+
  geom_vline(xintercept = which.max(storeK$GMean),lty = 2,col="red")
knnDemoPlot

```


##Demo KNN

```{r create example sets, include=FALSE}
numb<-nrow(knnDemo)
#Set a seed for sampling
set.seed(363663)
#Subset a small portion of the full data set for demonstration purpose 
DemoSubsetRow<-sample(1:numb,numb*0.05,replace = FALSE)
SetForDemo<-knnDemo[DemoSubsetRow,]  
```

```{r knn true demo visuals}
example_data<-data.frame("Age"= rep(NA,1),"Height"=rep(NA,1))
example_data$Age[1]<-8
example_data$Height[1]<-100

SetForDemo<-SetForDemo%>%
  mutate(dengue_status=as.character(dengue_status))
#demoTest<-example_data%>%
 # mutate(dengue_status_test=as.character(dengue_status))
ggplot()+
  geom_point(data = SetForDemo,aes(x=Age,y=Height,col=dengue_status),size=2,alpha=0.9,shape=19)+
  geom_point(data =example_data,aes(x=Age,y=Height),size=2,alpha=0.9,shape=19)+
  geom_segment(aes(x=8,xend=8,
                   y=100,yend=112.5))+
  geom_segment(aes(x=8,xend=7,
                   y=100,yend=112.5))+
  geom_segment(aes(x=8,xend=7,
                   y=100,yend=114))+
  geom_segment(aes(x=8,xend=9,
                   y=100,yend=117))+
  
  ggtitle("Figure 1: How KNN Works")
  
  
```

Here we have a data point (a boy with height of 100cm and age of 8 year old, shown as the black dot on the graph above).For instance, if we set our K for k-Nearest Neighbor Algorithm(KNN) equals to 4, the KNN Algorithm finds the dengue status of the 3 nearest data point respect to this data point on the graph(as shown in graph with black lines connect to the boy's data point) and check the dengue status of these 4 data points.The KNN then assigns the dengue status of the majority of these 4 data points as the predicted status of the boy. In this case, the dengue status of 3 out of 4 nearest data point to this boy's data point were shown as "1", indicating these three patients has dengue fever, therefore the boy was predicted to have dengue fever.



```{r creating actual KNN set and mutate logical var, include=FALSE}
knnActual<-dengue
```


### Split data in to 20% testing and 80% training 

```{r 80-20 split}
q<-nrow(knnActual)
#Set a seed for sampling
set.seed(363663)
manually_select<-sample(1:q,q*0.8,replace = "FALSE")
manual_train<-dengue[manually_select,]
manual_test<-dengue[-manually_select,]
storage8020<-data.frame("Predictions" = rep(NA,1146))
#Find rows with positive and negative result
    true1_split<-which(manual_test$Y == 1)
    true0_split<-which(manual_test$Y == 0)
    
#Compute the amount of rows corresponding to each result
    ntrue1_split<-length(true1_split)
    ntrue0_split<-length(true0_split)
    
storage_split<-data.frame("K"= rep(NA,50),"GeoMean"= rep(NA,50))

for(t in 1:50){
  storage_split$K[t]<-t
  k_preds_split<-knn(manual_train[,c(2,3,8,9)],manual_test[,c(2,3,8,9)],k=t,cl=manual_train$Y)
  storage8020$Predictions<-as.numeric(as.character(k_preds_split))
  
  sensitivity_split<-sum(k_preds_split[true1_split] == 1)/ntrue1_split

  specificity_split<-sum(k_preds_split[true0_split] == 0)/ntrue0_split
  storage_split$GeoMean[t]<-sqrt(sensitivity_split*specificity_split)

  

}
  
```


We utilized two approaches to predict the dengue status through KNN: (1) Creating a 20% test, 80% training split of the data and use it to for the KNN.(2)Using LOOCV or k-fold Cross Validation. For the first approach, 80% of the 5726 data entries were used for training purpose, and 20% of the data set were used to test the efficacy of the prediction. For the second approach, k-fold Cross Validation was selected, since the relative large size of the data set would lead to a executing time of the prediction. Only Age, DayDisease, Temp and BMI were kept for KNN since KNN can not accept categorical features. Height and Weight were excluded since BMI was calculated from them.


```{r 80-20 plot, echo=FALSE}
#Create plot to find the best k for KNN 
knnSplit<-ggplot(storage_split,aes(K,GeoMean))+
  geom_line()+
  labs(caption = paste("Geometric Mean, ReD Line at K=", which.max(storage_split$GeoMean)),title = "Figure 2: GeoMean Graph of KNN with Best k Shown
  (80% Training Data and 20% Testing Data)",y = " ")+
  geom_vline(xintercept = which.max(storage_split$GeoMean),lty = 2,col="red")
knnSplit
```

For the first approach, we determined the best k value for KNN through running KNN with k from 1-50 and calculating the geometric mean of the sensitivity and specificity of prediction under each k value. We found k=46 to be the best k value since the prediction under k=46 provide the most balanced sensitivity(true positive rate) and specificity(true negative rate) in the prediction of dengue fever status.

```{r}
knitr::kable(table(storage8020$Predictions, manual_test$Y), caption= "Table 1: Confusion Matrix of Dengue Status Prediction(80% Training & 20% Testing ) ", col = c("Not Dengue", "Dengue"))
```

From the 80% training, 20% testing approach, we have:
$$Sensitivity = \frac{711}{711+87}\approx 0.891$$
$$Specificity = \frac{122}{122+226}\approx 0.351$$
$$Accuracy = \frac{122+711}{122+711+226+87}\approx 0.727$$


```{r 10 fold CV determine best k }
n<-nrow(knnActual)
nk<-50
storage<-data.frame("K"=rep(NA,nk),
                    "Sensitivity" = rep(NA,nk),
                    "Specificity" = rep(NA,nk),
                    "GeoMean" = rep(NA,nk))

#Set a seed for sampling and create a pool and set fold for K fold CV
set.seed(363663)
pool<-rep(1:10,573)

fold<-sample(pool,5726,replace = FALSE)

#Outer loop for increment k 
for(k in 1:nk){
  storage$K[k]<-k
  
  storage_inner<-data.frame("YHat" = rep(NA,n))
  #Inner loop for 10 fold CV
  for(i in 1:10){
    #Find data in each fold
    infold<-which(fold == i)
    
    #Create training and testing sets
    Train_Dengue<-knnActual[-infold,]
    Test_Dengue<-knnActual[infold,]
    
    #Run Knn
    k_preds<-knn(Train_Dengue[,c(2,3,8,9)],Test_Dengue[,c(2,3,8,9)],k=k,cl=Train_Dengue$Y)
    
    #store results
    storage_inner$YHat[infold]<-as.numeric(as.character(k_preds))
    
    #Find rows with positive and negative result
    true1K<-which(Test_Dengue$Y == 1)
    true0K<-which(Test_Dengue$Y == 0)
    
    #Compute the amount of rows corresponding to each result
    ntrue1K<-length(true1K)
    ntrue0K<-length(true0K)
#Compute sensitivity and specificity, as well as GeoMetric Mean for determining the best value for k
sensitivity<-sum(k_preds[true1K] == 1)/ntrue1K
storage$Sensitivity[k]<-sensitivity
specificity<-sum(k_preds[true0K] == 0)/ntrue0K
storage$Specificity[k]<-specificity
storage$GeoMean[k]<-sqrt(sensitivity*specificity)
  }
}




```

$$Sensitivity = \frac{3608}{3608+420}\approx 0.895$$

$$Specificity = \frac{544}{554+1154}\approx 0.320$$

$$Accuracy = \frac{544+3608}{5726}\approx 0.725$$



```{r find best k}
#Create plot to find the best k for KNN 
knnActualPlot<-ggplot(storage,aes(K,GeoMean))+
  geom_line()+
  labs(caption = paste("Geometric Mean, ReD Line at K=", which.max(storage$GeoMean)),title = "Figure 3: GeoMean Graph of KNN with Best k Shown
  (10 Fold Cross Validation)",y = " ")+
  geom_vline(xintercept = which.max(storage$GeoMean),lty = 2,col="red")
#Show Plot
knnActualPlot
```

For the k fold Cross Validation, we found k=35 to be the best k value for 10-fold Cross Validation because it provides the largest geometric mean among k values from 1-50, indicating the most ideal balance between the sensitivity and specificity of the prediction.

```{r}
knitr::kable(table(storage_inner$YHat, dengue$Y), caption= "Table 2: Confusion Matrix of Predicting Dengue Status (10 Fold CV & KNN with K = 35)  ", col = c("Not Dengue", "Dengue"))  
```

We have
$$Sensitivity = \frac{3608}{3608+420}\approx 0.895$$

$$Specificity = \frac{544}{554+1154}\approx 0.320$$

$$Accuracy = \frac{544+3608}{5726}\approx 0.725$$

```{r k equals 35, eval=FALSE, include=FALSE}
#Set a seed for sampling
set.seed(363663)
#create storage for prediction result
storage_inner_trial<-data.frame("pred"=rep(NA,5726))
fold35<-sample(pool,5726,replace = FALSE)

#run 10 fold knn with best k value = 35
 for(i in 1:10){
   
   #Find data in each fold
    infold2<-which(fold35 == i)
    
    #Creating Training and testing set
    Train_Dengue<-knnActual[-infold2,]
    Test_Dengue<-knnActual[infold2,]
    #Run knn
    k_preds_best<-knn(Train_Dengue[,c(2,3,8,9)],Test_Dengue[,c(2,3,8,9)],k=35,cl=Train_Dengue$Y)
    trueBestk1<-which(Test_Dengue$Y==1)
    trueBestk0<-which(Test_Dengue$Y==0)
    
    ntrueBest1<-length(trueBestk1)
    ntrueBest0<-length(trueBestk0)
    
    #Compute sensitivity and specificity,formating them
    sensitivity35<-sum(k_preds_best[trueBestk1] == 1)/ntrueBest1
    print(ntrueBest1)
    specificit35<-sum(k_preds_best[trueBestk0] == 0)/ntrueBest0

    #store result
    storage_inner_trial$pred[infold2]<-as.numeric(as.character(k_preds_best))
    
    

    
 }



#Create confusion Matrix for the prediction with full dengue set

```






## Section 3 Logistic Regression 

```{r empirical logit function, include=FALSE}
#Load function for empirical logit plot
emplogitPlot <- function(x, y, binsize = NULL, ci = FALSE, probit = FALSE,
prob = FALSE, main = NULL, xlab = "", ylab = "", lowess.in = FALSE){
  
  if(class(y) =="character"){
   y <- as.numeric(as.factor(y))-1
   }
  
  if (length(x) != length(y))
    stop("x and y lengths differ")
  if (any(y < 0 | y > 1))
    stop("y not between 0 and 1")
  if (length(x) < 100 & is.null(binsize))
    stop("Less than 100 observations: specify binsize manually")
  
  if (is.null(binsize)) binsize = min(round(length(x)/10), 50)
  
  if (probit){
    link = qnorm
    if (is.null(main)) main = "Empirical probits"
  } else {
    link = function(x) log(x/(1-x))
    if (is.null(main)) main = "Empirical logits"
  }
  
  sort = order(x)
  x = x[sort]
  y = y[sort]
  a = seq(1, length(x), by=binsize)
  b = c(a[-1] - 1, length(x))
  
  prob = xmean = ns = rep(0, length(a)) # ns is for CIs
  for (i in 1:length(a)){
    range = (a[i]):(b[i])
    prob[i] = mean(y[range])
    xmean[i] = mean(x[range])
    ns[i] = b[i] - a[i] + 1 # for CI 
  }
  
  extreme = (prob == 1 | prob == 0)
  prob[prob == 0] = min(prob[!extreme])
  prob[prob == 1] = max(prob[!extreme])
  
  g = link(prob) # logits (or probits if probit == TRUE)
  
  linear.fit = lm(g[!extreme] ~ xmean[!extreme])
  b0 = linear.fit$coef[1]
  b1 = linear.fit$coef[2]
  
  loess.fit = loess(g[!extreme] ~ xmean[!extreme])
  
  plot(xmean, g, main=main, xlab=xlab, ylab=ylab)
  abline(b0,b1)
  if(lowess.in ==TRUE){
  lines(loess.fit$x, loess.fit$fitted, lwd=2, lty=2)
  }
}
```

#EDA 


##Checking Conditions and Interactions


```{r creating dedicated data set}
#Create a data set dedicated for checking condition and EDA
dengueForCheck<-knnActual
```


```{r empirical logit plot}
#Display 4 graphs together
op<-par(mfrow=c(2,2))

#Checking linear relationship between log odds that Y=1 and four numeric Xi
logit_DayDisease<-emplogitPlot(x=dengueForCheck$DayDisease,y=dengueForCheck$Y,xlab = "Day Diseases",ylab = "Log Odds of Dengue",main = "Figure 3.1 Empirical Logit Plot Day Disease")

logit_Temp<-emplogitPlot(x=dengueForCheck$Temp,y=dengueForCheck$Y,xlab = "Temperature",ylab = "Log Odds of Dengue",main = "Figure 3.2 Empirical Logit Plot Temperature")

logit_Age<-emplogitPlot(x=dengueForCheck$Age,y=dengueForCheck$Y,xlab = "Age",ylab = "Log Odds of Dengue",main = "Figure 3.3 Empirical Logit Plot Age")

logit_BMI<-emplogitPlot(x=dengueForCheck$BMI,y=dengueForCheck$Y,xlab = "BMI",ylab = "Log Odds of Dengue",main = "Figure 3.4 Empirical Logit Plot BMI")

```



```{r box plot sex and temperature}
#Convert sex from a numeric feature to categorical
dengueForCheck$Sex=as.factor(dengueForCheck$Sex)
BoxDengue<-ggplot(data=dengueForCheck,aes(x=Sex,y=Temp,fill=Temp))+
  theme_bw()+
  geom_boxplot()
BoxDengue
```

No significant difference on body temperature across sex was found. 

```{r scatter plot to see relationship between age and dengue status}
knnActual%>%
  mutate(dengue_status=as.factor(Y))%>%
  ggplot(aes(x=Age,y=Height,color=dengue_status))+
  geom_point()+
  ggtitle("Figure 3.6: Age vs Height Plot Respect to Dengue Status")
```


```{r exploring correlation}
#Checking correlations between variable to determine multicolinearity
correlation<-cor(dengueForCheck[,c(2,8:11)])
correlation_tr<-round(correlation,3)
knitr::kable(correlation_tr)
```

We suspected that multicolinearity exist between height and weight, as well as Age and height/weight. We also suspect multicolinearity appear between BMI and Height/Weight, since BMI was calculated from these two features. An interaction term between age and weight will be added to deal with the multicolinearity Additionally, vomiting and rash are two of the common symptoms of dengue fever.Thus, age, BMI,weight, temperature will be numeric features selected for the logistic regression model.Sex, Rash and Vomiting will be the categorical features selected for the logistic regression. 


```{r train logistic with 10-fold CV}
#Set seed for sampling
set.seed(363663)
n<-nrow(knnActual)
storage_logistic<-data.frame("Predicted" = rep(NA,5726))

#Creating pools and set fold for K fold CV
logisticpool<-rep(1:10,573)

#Sampling for the pool, assign rows with fold number 
logisticfold<-sample(logisticpool,5726,replace = FALSE)

#Subset a data set for training logistic model
logisticTotal<-knnActual[,c(1,2,4,8,9,11,12,13,14,15)]

#Create for loop to train the model 
for(i in 1:10){
  inlogistic<-which(logisticfold==i)
  logisticTrain<-logisticTotal[-inlogistic,]
  logisticTest<-logisticTotal[inlogistic,]
  
  modelTrained<-glm(data=logisticTrain,family = binomial,Y~Sex+Vomiting+Temp+Age*Weight+BMI+Rash+Age+Weight)
  probabilities<-predict(modelTrained,newdata=logisticTest)
  predicted.Y<-ifelse(probabilities>0.5,"1","0")
  storage_logistic$Predicted[inlogistic]<-as.numeric(as.character(predicted.Y))
  
}

modelTotal<-glm(data=dengue,family = binomial,Y~Sex+Vomiting+Temp+Age*Weight+BMI+Rash+Age+Weight)
storage_modelTotal<-data.frame("Prediction"= rep(NA,5726))
ModelTotalprobabilities<-predict(modelTotal,newdata=dengue)
predicted.YTotal<-ifelse(ModelTotalprobabilities>0.5,"1","0")
storage_modelTotal$Prediction<-as.numeric(as.character(predicted.YTotal))
knitr::kable(table(storage_logistic$Predicted, dengue$Y), caption= "Table x: Confusion Matrix of Predicting Dengue Status (Logistic Regerssion & 10 Fold CV) ", col = c("Not Dengue", "Dengue"))  
```

$$Sensitivity = \frac{3138}{890+3138}\approx 0.779$$

$$Specificity = \frac{939}{759+939}\approx 0.553 $$

$$Accuracy = \frac{939+3138}{5726} \approx 0.712$$

# Formatted Confusion Matrix of Prediction with logistic regression

```{r}
knitr::kable(table(storage_logistic$Predicted, knnActual$Y), caption= "Table 3: Confusion Matrix of Predicting Dengue Status (Logistic Regerssion & 10 Fold cv)  ", col = c("Not Dengue", "Dengue"))  
```


$$Sensitivity = \frac{3138}{759+3138}\approx 0.805$$

$$Specificity = \frac{939}{759+939}\approx 0.553$$

```{r 80 20 Split model for logistic, eval=FALSE, include=FALSE}
logisticModel8020<-glm(data=manual_train,family = binomial,Y~Sex+Vomiting+Temp+Age*Weight+BMI+Rash)
logisticModel8020
```


```{r 8020 logistic, eval=FALSE, include=FALSE}
storageLogi8020<-data.frame("Prediction"= rep(NA,5726))
LogiProbabilities<-predict(logisticModel8020,newdata=knnActual)
predicted.Y8020<-ifelse(LogiProbabilities>0.5,"1","0")
storageLogi8020$Prediction<-as.numeric(as.character(predicted.Y8020))
knitr::kable(table(storageLogi8020$Prediction, dengue$Y), caption= "Table x: Confusion Matrix of Predicting Dengue Status (Logistic Regerssion & 80% Training 20% Testing)  ", col = c("Not Dengue", "Dengue"))  
```


## Section 4 Discussion

|             | KNN   | Logistic | Winner   |   |
|-------------|-------|----------|----------|---|
| Sensitivity | 0.272 | 0.02     | KNN      |   |
| Specificity | 0.964 | 1        | Logistic |   |
| Accuracy    | 0.893 | 0.9      | Logistic |   |




## Bibliography
CDC. (2021, September 20). Symptoms and treatment. Centers for Disease Control and Prevention. Retrieved September 28, 2022, from https://www.cdc.gov/dengue/symptoms/index.html\

